\SW: Crowd-sourcing the Discovery of Gravitational Lenses

### 2014-09-02 doc/sw-system.tex
Short piece of text about the P vs C plot, needs numbers from @cpadavis #65

This paper is organised as follows.  In \Sref{sec:design} we introduce the \sw
classification interface and the volunteers who make up the \sw
collaboration,  explain how we use training images, and describe our two stage
candidate selection strategy. We then briefly introduce, in \Sref{sec:data}
the particular dataset used in the first experimental tests of the \sw system,
and how we prepared the images prior to displaying them in the web interface.
In \Sref{sec:swap} we describe our methodology for interpreting the
classifications made by the volunteers, and then present the results of system
performance tests  made on the training images in \Sref{sec:results}. % In
this section, we also investigate the properties of the crowd, to % understand
where the information is coming from. We discuss the implications of our
results for future lens searches in \Sref{sec:discuss} and draw conclusions in
\Sref{sec:conclude}.

@@

We now quantify the performance of the \sw system in terms of the recovery of
the training set images. \Fref{fig:results:sample:roc} shows receiver
operating characteristic (ROC) curves for CFHTLS stage 1 and stage 2. These
plots show the true positive rate (TPR), the number of sims correctly detected
divided by the total number of sims in the training set, and the false
positive rate (FPR), the number of duds incorrectly detected  divided by total
number of duds in the training set, both for a given sample of detections defined by
a particular probability threshold, which varies along the curves.  In both
stages, these curves show that true positive rates of  around 90\% were
achieved, at very low false positive rates. For comparison we show the results
of an analysis where the classifications of  training images were ignored, and
none of the agents allowed to learn: they were instead assigned initial values
of their confusion matrices of 0.75 for each element, which then remained
constant. This emulates a very simple unweighted voting scheme, where all
classifications are treated equally. In this case, the TPR remains under 80\%
in stage 1 and 60\% in stage 2, showing the benefit of including training
images and allowing the agents to learn. The choice of initial confusion
matrix is not very important: the same 0.75 initial values applied to normal,
learning agents results in a slightly lower TPR than the default case,
indicating that initializing the system less conservatively results in
slightly worse performance.

The dot-dashed curves show the impact of the offline analysis. At stage 1 the
results are very similar to the online version that was actually run (solid
line), at stage 2 there seems to be some benefit to doing the analysis in this
way. Remarkably, over 85\% TPR is achieved at zero FPR. Alternatively, if one
is willing to accept a false positive rate of 5\%, the true positive rate
rises to over 95\%, showing that some of the sims that were missed in the
online analysis can be recovered by doing the analysis offline. The same is
true at stage 1, but to a much lower degree.

@@

Adopting the online stage 1 analysis, and the offline stage 2 analysis, we show
in \Fref{fig:results:sample:CP} a plot of completeness versus purity in the two
stages. As in \Fref{fig:results:sample:roc}, the detection threshold varies
along the curves. Completeness is defined as the number of correctly detected
sims divided by the total number of sims in the training set, while purity is
the number of correct detections divided by the total number of
detections.\footnote{The completeness is equivalent to the TPR and is also
known elsewhere as the `recall'. The purity is also known as the `precision'.}
If the training set samples from the same systems as the test set, then the
completeness of the training set will be equivalent to the completeness of the
test set. The purity depends on the proportion of sims to duds, and so the
purity of the test set must be approximated by rescaling the training set to
the expected proportion of lens systems to not-lens systems in the survey.
First we compute the expected number of false positives by multiplying the FPR
by the expected number of non-lenses in the survey. Then we multiply the TPR by
the expected number of lenses in the survey, to get the expected number of true
positives. The sum of the true positives and the false positives gives the
expected sample size; dividing the expected number of true positives by this
sample size gives the purity. Note that the completeness is invariant to this
transformation.

\todo{Chris}{Write analysis of completeness purity.}

>>>>

Our basic method in this paper is to analyze the performance of the \sw system
on a ``training set'' of simulated lenses and known non-lenses. This allows us
to estimate completeness and purity with respect to gravitational lenses that
have the same properties of the training set. In the companion paper we carry
out similar tests but with a sample of known (real) lenses {\it in situ}.

This paper is organised as follows.  In \Sref{sec:design} we introduce the \sw
classification interface and the volunteers who make up the \sw
collaboration,  explain how we use training images, and describe our two stage
candidate selection strategy. We then briefly introduce, in \Sref{sec:data}
the particular dataset used in the first experimental tests of the \sw system,
and how we prepared the images prior to displaying them in the web interface.
In \Sref{sec:swap} we describe our methodology for interpreting the
classifications made by the volunteers, and then present the results of system
performance tests  made on the training images in \Sref{sec:results}.  We
discuss the implications of our results for future lens searches in
\Sref{sec:discuss} and draw conclusions in \Sref{sec:conclude}.

@@

We now quantify the performance of the \sw system in terms of the recovery of
the training set images. \Fref{fig:results:sample:roc} shows receiver
operating characteristic (ROC) curves for CFHTLS stage 1 and stage 2. These
plots show the true positive rate (TPR), the number of sims correctly detected
divided by the total number of sims in the training set, and the false
positive rate (FPR), the number of duds incorrectly detected  divided by total
number of duds in the training set, both for a given sample of detections defined by
a particular probability threshold, which varies along the curves.  In both
stages, these curves show that true positive rates of  around 90\% were
achieved, at very low false positive rates. For comparison we show the results
of an analysis where the classifications of  training images were ignored, and
none of the agents allowed to learn: they were instead assigned initial values
of their confusion matrices of 0.75 for each element, which then remained
constant. This emulates a very simple unweighted voting scheme, where all
classifications are treated equally. In this case, the TPR remains under 80\%
in stage 1 and 60\% in stage 2, thus quantifying 
the benefit of including training
images and allowing the agents to learn. The choice of initial confusion
matrix is not very important: the same 0.75 initial values applied to normal,
learning agents results in a slightly lower TPR than the default case,
indicating that initializing the system less conservatively results in
slightly worse performance.

The dot-dashed curves show the impact of the offline analysis. At stage 1 the
results are very similar to the online version that was actually run (solid
line). However, at stage 2 there seems to be somewhat greater benefit to doing
the analysis in this way: over 85\% TPR is achieved at zero FPR.
Alternatively, if one is willing to accept a false positive rate of 5\%, the
true positive rate rises to over 95\%, showing that some of the sims that were
missed in the online analysis can be recovered by doing the analysis offline.
The same is true at stage 1, but to a lower degree.

@@

Adopting the online stage 1 analysis, and the offline stage 2 analysis, we
show in \Fref{fig:results:sample:CP} a plot of the more familiar (to
astronomers) quantities completeness versus purity in
the two stages. As in \Fref{fig:results:sample:roc}, the detection threshold
varies along the curves. Completeness is defined as the number of correctly
detected sims divided by the total number of sims in the training set, while
purity is the number of correct detections divided by the total number of
detections.\footnote{The completeness is equivalent to the TPR and is also
known elsewhere as the ``recall.''  The purity is also known as the
``precision.''} If the training set samples from the same systems as the test
set, then the completeness of the training set will be equivalent to the
completeness of the test set. The purity depends on the proportion of sims to
duds, and so the purity of the test set must be approximated by rescaling the
training set to the expected proportion of lens systems to not-lens systems in
the survey. First we compute the expected number of false positives by
multiplying the FPR by the expected number of non-lenses in the survey. Then
we multiply the TPR by the expected number of lenses in the survey, to get the
expected number of true positives. The sum of the true positives and the false
positives gives the expected sample size; dividing the expected number of true
positives by this sample size gives the purity. Note that the completeness is
invariant to this transformation.

The stage 1 curves are truncated by the retirement of subjects in this phase, 
which sets the minimum size of this sample. We see from the solid blue curve
that over 90\% completeness was reached, albeit in a sample with around 20\%
purity. The lensed quasar part of the training set yielded the highest
completeness, suggesting that these were the easiest sims to spot. The lensed
galaxies were recovered at understandably lowest completeness, although the
wider image separation  lensing clusters were not very different. At stage 2,
where no retirement was carried out, it was possible to reach 100\% purity:
indeed, the knee of the curve is at just under 90\% completeness. However, the
purity decreases rapidly if higher completeness than this is sought.
Interestingly, the lensing clusters can be seen to pose problems for the
system, with its curve turning over at just 60\% completeness.

The optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > X.XX$. At 100\% purity
and 9X\% completeness, it would have contained around N lens candidates.
